---
title: "STA 3241 Final Project"
output:
  html_document:
    toc: true
    toc_depth: '4'
    df_print: paged
    toc_float: true
  html_notebook:
    number_sections: no
    toc: true
    toc_depth: 4
    toc_float: true
  pdf_document:
    toc: true
    toc_depth: '4'
---
# ***Library Loading:***
```{r}
library(tidyverse)
library(dplyr)
library(openxlsx)
library(tidytext)
library(forcats)
library(lubridate)
library(data.table)
```

# ***Data Frame Creation:***
```{r}
ufo <- data.table::fread("ufo-complete-geocoded-time-standardized.csv")
ufo$Seconds <- as.double(ufo$Seconds)
ufo$Shape <- as.factor(ufo$Shape)
ufo$State_Province <- as.factor(ufo$State_Province)
ufo$Country <-  as.factor(ufo$Country)
ufo$Latitude <- as.double(ufo$Latitude)
ufo$Longitude <- as.double(ufo$Longitude)
ufo$City <- as.factor(ufo$City)
ufo$Date <- as.Date(ufo$Date, format = "%m/%d/%Y")
ufo$Time <- as.ITime(ufo$Time, tz = "GMT")
ufo
```


# ***Observations by Country:***
```{r}
ufo %>%
  group_by(Country)%>%
  filter(n() >= 200)%>%
  ggplot() +
  geom_bar(aes(x = Country, fill = Country))+ 
  coord_flip()+
  labs(title = "Reports by Country", 
       y = "Number of Specific Reports")+
  theme(legend.position ="none")
```

As we can see, nearly all of our data is showing reports from within the United States. 
Canada takes second place, followed by England. 
It's noteworthy that Australia, India, and Mexico are displayed on this graph as there are over 600 unique levels. 

## **Which States from the USA?**
```{r}
ufo %>%
  filter(Country == "United States")%>%
  group_by(State_Province)%>%
  filter(n() >= 1000)%>%
  ggplot()+
  geom_bar(aes(x = State_Province, fill = State_Province))+ 
  coord_flip()+
  labs(title = "Reports by State", 
       y = "Number of Specific Reports")+
  theme(legend.position ="none")#+
  #facet_wrap(~ State_Province, scales ="free_y")
```

Will work on the visualization, but as we can see, California has some of the highest reports out there. 

## **What Time are these normally viewed at?**
```{r}
ufo %>%
  filter(Country == "United States")%>%
  group_by(State_Province)%>%
  filter(n() >= 1000)%>%
  ggplot()+
  geom_col(aes(x = State_Province, y = Time, fill = State_Province))+ 
  coord_flip()+
  labs(title = "Viewing Time by State", 
       y = "Time")+
    theme(legend.position ="none")
```


# ***Shapes Reported within the U.S.***
```{r}
ufo %>%
  filter(Country == "United States")%>%
  group_by(Shape)%>%
  filter(n() >= 1000)%>%
  ggplot()+
  geom_bar(aes(x = Shape, fill = Shape))+ 
  coord_flip()+
  labs(title = "Shape Tendencies", 
       y = "Number of Specific Reports")+
    theme(legend.position ="none")
```

Light is the most frequent response. 
However, this is followed up by circularity and triangular shapes. 



# ***Oh, Florida:***
```{r}
ufo %>%
  filter(Country == "United States")%>%
  filter(State_Province == "fl")%>%
  group_by(Shape)%>%
  filter(n() >= 100)%>%
  ggplot()+
  geom_bar(aes(x = Shape, fill = Shape))+ 
  coord_flip()+
  labs(title = "Shape Tendencies for Florida", 
       y = "Number of Specific Reports")+
    theme(legend.position ="none")
```



## Locations of Reports in FL
```{r}
ufo %>%
  filter(Country == "United States")%>%
  filter(State_Province == "fl")%>%
  filter(Latitude <= 140)%>%
  filter(Longitude >= -100)%>%
  ggplot()+
  geom_point(aes(x = Latitude, y = Longitude, color = Shape))+ 
  labs(title = "Shape Tendencies for Florida")+
    theme(legend.position ="none")
```

